---
title: "script"
author: "Stephane Le Vu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{script}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Pipeline for analysis of UK Drug resistance database

- Raw data from UK Drug Resistance database and Public Health England collaboration received on 13 September 2016.
- Subtype specific (A, B, C, CRF02AG) HIV1 global sequences (one per patient) from Los Alamos National Laboratory, were downloaded on 7 December 2016, with position from 2253 to 3798, including HXB2. Sequences originated from the UK were removed from these.
- In the package data are already pre-processed with some recoding and can be obtained from:

```{r see_data}
library(tenbrit)
d <- data(package = "tenbrit")
d$results[,3:4]
```
- And documentation for processed data, as well as initial raw file (in `References`), is available from:
```{r see_doc, eval = FALSE}
help(resi)
```

All subsequent code comes from `garel` package
```{r lib}
library(garel)
rerun <- TRUE
```

### Load UK and LANL sequences
- I stratify multi sequence alignments of UK sequences in `s950` into 8 (transmission mode $\times$ subtype) groups (creates `data` folder)

```{r stratify_sequences, eval = rerun}
stratify_msa()
```

- And I copy the global sequences MSAs also here
```{r copy_lanl, eval = rerun}
lanl_msa()
```

- Output are here:
```{r, eval = TRUE}
uk_files <- list.files("data", "^MSM_.*\\.(rda|rds)$", full.names = TRUE)
uk_files
lanl_files <- list.files("data", "^db_.*\\.(rda|rds)$", full.names = TRUE)
lanl_files
```
### make a smaller msa
```{r downsample, eval = rerun}
seqs <- readRDS("./data/MSM_CRF02AG.rds")
set.seed(10)
saveRDS(seqs[sample(length(seqs),size = 100)], file = "data/MSM_CRF02AG_small.rds")
uk_files <- list.files("data", "^MSM_.*\\.(rda|rds)$", full.names = TRUE)
uk_files
```

### Make BLAST databases
- With a pipeline that
    - splits MSA in small chunks (100 sequences each by default)
    - removes drug resistance mutation sites
    - merges back the splits
    - makes a BLAST database in `blastDB` folder
- Here I run the non-HPC version on subtype CRF02AG
```{r make_blastdb, eval = rerun}
pipeline_blastdb(lanl_files[4], verbose = FALSE)
list.files("./blastDB")
```

### Process UK sequences BLAST UK sequences and add aligned hits
- With pipeline that
    - splits MSA in smaller chunks (in new folder `*_splits`)
    - removes drug resistance mutation sites
    - BLAST UK sequences against corresponding global database
    - merges back the splits
    - add aligned hits
    - and add ref sequences as outgroup for rooting
- It generates
    - a list of unique BLAST hits `*_nodrm_unique_blast_hits.rds`
    - an alignement ready for tree estimation in the form `*_nodrm_ref_og.fas`
```{r process_UK_seq, eval = rerun}
pipeline_msa(path.msa = uk_files[4], verbose = FALSE)
(LSTFAS <- list.files(pattern = "_nodrm_ref_og.fas$", recursive = TRUE, full.names=TRUE))
```



### Maximum likelihood trees
- Make first tree (here on local machine)
```{r tree0, eval = rerun}
pipeline_tree0(pathmsa=LSTFAS[1])
```

Have a look at the tree, with UK sequences in black and global sequences in red
```{r plot_tree0}
TREE <- "data/MLTREE0/ExaML_result.MSM_CRF02AG.finaltree.000"
tr <- ape::read.tree(TREE)
plot_tree_col_og(tr)
```

```{r bs_trees}
FLTDMSAS <- list.files(pattern = "_fltd.fas$", recursive = TRUE, full.names=TRUE)

pipeline_bstrees(
  pathmsa = FLTDMSAS[1],
  dirbstree = "data/BSTREES",
  BSD=0, BSE=2, Q = NA)
```

plot

```{r plot2}
POSTLSD <- list.files(pattern = "ExaML_result.*post_lsd\\.\\d{3}_lsd\\.rds$", recursive = TRUE)
plot_all_lsd_tree(POSTLSD)

script <- lapply(POSTLSD, function(x){
      Sys.sleep(.1)
      hpc.cmd.funr(f.name = "extract_clades",
                   f.arg = paste0("lsd_results=", x),
 hpc.time = "#PBS -l walltime=00:59:00",
                   submit = TRUE,
                   hpc.q = NA)
    })

path_clades <- list.files(path = "data/CLADES", pattern = "clades", full.names = TRUE)

iter <- 1:3
dummy <- lapply(path_clades, function(x) {
script <- sapply(iter, function(i){
  Sys.sleep(.1)
  hpc.cmd.funr(f.name = "get_ip",
               f.arg = paste0("pathclade=", x),
               hpc.time = "#PBS -l walltime=00:59:00",
               submit = TRUE,
               hpc.q = NA, 
               verbose = TRUE)
})
})

```

Process infector probs

```{r}
IP <- list.files(path ="data/IPS", pattern = "ip.*\\.rds$", full.names = TRUE)
list_mean_ips_per_bs <- mean_ip_per_bs(pathip=IP)
```

```{r}
data("df", package = "tenbrit", envir = environment())
sub_df_msm <- df[ grepl("MSM", df$group), ]
```

```{r}
list_ip_by_age_bs <- lapply(c(age_cl4 = 'age_cl4'),
                           function(v) {
                             lapply(list_mean_ips_per_bs, function(g){
                               lapply(g, function(x) ip_by_var(x, sub_df_msm, v))
                             }) 
                           })

```

```{r}
var_order <- list(
age_cl4 = as.character(quantile(sub_df_msm$ageres, na.rm = T)[-1])
)
vars <- "age_cl4"  
system.time(
    l_all_mixmat_bs <- lapply(1:length(vars),
                           function(i){
                             l <- lapply(list_ip_by_age_bs[[vars[i]]], function(bs){
                               lapply(bs, edgelist_to_mixingmatrix)
                             })
                             return(l)
                           }) # 6s
  )
  names(l_all_mixmat_bs) <- vars
  ##- reorder all mixing matrix
  l_mixmat <- lapply(setNames(names(l_all_mixmat_bs),names(l_all_mixmat_bs)), 
                     function(var){
    lapply(l_all_mixmat_bs[[var]], function(bs){
      lapply(bs, function(mat) rorder(m = mat, o = var_order[[var]], normalize = TRUE))
    })
  })
```
```{r}
##- 4 matrices (a, b, c ,d) as (M, R, D, A), plus r
l4 <- lapply(l_mixmat[[1]], function(x){
  lapply(x, function(m){
    list(a= m,
         b = m / rowSums(m)[row(m)],
         c = m / colSums(m)[col(m)],
         d = mat2assortmat(m),
         r = newman_r(m))
  })
})
ll4 <- lapply(l4, ukdrdb:::revert_list)
```

```{r}
ci_matrix <- function(lsmat, k = c(.025, 0.975)){
    d <- dim(lsmat[[2]])
    if(!all(sapply(lsmat, dim) == d)) stop("must have matrices with same dim")
    rn <- rownames(lsmat[[1]])
    cn <- colnames(lsmat[[1]])
    a <- sapply(lsmat, as.vector)
    qs <- lapply(setNames(k, k), function(q) {
      apply(a, 1 , function(x) quantile(x, probs = q))
    })
    ms <- lapply(qs, function(x){
      m <- matrix(x, nrow = d, byrow = F)
      rownames(m) <- rn
      colnames(m) <- cn
      m
    })  
    ms
  }
  
  # ci_matrix(lsmat = ll4[['MSM_A']][['a']])
  lll4 <- lapply(ll4, function(l) lapply(l[1:4], ci_matrix))
```

```{r}
array_all_mat <- lapply(ll4, function(st){
  lapply(st, function(lsmat){ # doesn't include newman coef
    ## array of 100 m*m matrices
    ar <- sapply(lsmat, identity, simplify="array")
    if( is.null(dim(ar)) ){ # vector
      x <- ar
      c(mean = mean(x), median = median(x), 
        quantile(x, probs = c(.025, 0.975)), sd = sd(x))
    } else { # matrix
      arm <- apply(ar, 1:2, function(x){
        c(mean = mean(x), median = median(x), 
          quantile(x, probs = c(.025, 0.975)), sd = sd(x))
      }) ## beware: dim = c(1:4, dim(m))
      #dim(arm) # c(4,4,4)
      arm
    }
    #dim(ar) # c(4, 4, 100)
  })
})
array_all_mat$MSM_CRF02AG$a[1,,] # matrix of mean
array_all_mat$MSM_CRF02AG$a[3,,] # lo CI
array_all_mat$MSM_CRF02AG$a[4,,] # up CI
```

