---
title: "script"
author: "Stephane Le Vu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{script}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Pipeline for analysis of UK Drug resistance database

- Raw data from UK Drug Resistance database and Public Health England collaboration received on 13 September 2016.
- Subtype specific (A, B, C, CRF02AG) HIV1 global sequences (one per patient) from Los Alamos National Laboratory, were downloaded on 7 December 2016, with position from 2253 to 3798, including HXB2. Sequences originated from the UK were removed from these.
- In the package data are already pre-processed with some recoding and can be obtained from:

```{r see_data}
library(tenbrit)
d <- data(package = "tenbrit")
d$results[,3:4]
```
- And documentation for processed data, as well as initial raw file (in `References`), is available from:
```{r see_doc, eval = FALSE}
help(resi)
```

All subsequent code comes from `garel` package
```{r lib}
library(garel)
rerun <- FALSE
```

### Load UK and LANL sequences
- I stratify multi sequence alignments of UK sequences in `s950` into 8 (transmission mode $\times$ subtype) groups (creates `data` folder)

```{r stratify_sequences, eval = rerun}
stratify_msa()
```

- And I copy the global sequences MSAs also here
```{r copy_lanl, eval = rerun}
lanl_msa()
```

- Output are here:
```{r, eval = TRUE}
uk_files <- list.files("data", "^MSM_.*\\.(rda|rds)$", full.names = TRUE)
uk_files
lanl_files <- list.files("data", "^db_.*\\.(rda|rds)$", full.names = TRUE)
lanl_files
```
### make a smaller msa
```{r}
seqs <- readRDS("./data/MSM_CRF02AG.rds")
saveRDS(seqs[sample(length(seqs),size = 20)], file = "data/MSM_CRF02AG_small.rds")
uk_files <- list.files("data", "^MSM_.*\\.(rda|rds)$", full.names = TRUE)
uk_files
```

### Make BLAST databases
- With a pipeline that
    - splits MSA in small chunks (100 sequences each by default)
    - removes drug resistance mutation sites
    - merges back the splits
    - makes a BLAST database in `blastDB` folder
- Here I run the non-HPC version on subtype CRF02AG
```{r make_blastdb, eval = rerun}
pipeline_blastdb(lanl_files[4], verbose = FALSE)
list.files("./blastDB")
```

### Process UK sequences BLAST UK sequences and add aligned hits
- With pipeline that
    - splits MSA in smaller chunks (in new folder `*_splits`)
    - removes drug resistance mutation sites
    - BLAST UK sequences against corresponding global database
    - merges back the splits
    - add aligned hits
    - and add ref sequences as outgroup for rooting
- It generates
    - a list of unique BLAST hits `*_nodrm_unique_blast_hits.rds`
    - an alignement ready for tree estimation in the form `*_nodrm_ref_og.fas`
```{r process_UK_seq, eval = TRUE}
## CRF02AG only
pipeline_msa(path.msa = uk_files[4], verbose = FALSE)
(LSTFAS <- list.files(pattern = "_nodrm_ref_og.fas$", recursive = TRUE, full.names=TRUE))
```



### Maximum likelihood trees
- Make first tree (here on local machine)
```{r tree0, eval = rerun}
pipeline_tree0(pathmsa=LSTFAS[1])
```


